---
title: "NodeDelAdd"
output: html_notebook
---
```{r}
library(biclust)
yst <- read.csv("data/BicatYeast.csv")
```

```{r}
resBIC <- biclust(yst,method=BCCC(),delta=150,alpha=1.2,number=100)
```
Looks like this may be a problem due to OO facilities.
Looks like new version of biclust doesn't have same data set as book used.

Brief from  2008 has new name of dataset (but dimension of data is different)
I have biclust 2.0.3.1 5/19/2023

#Delta-Biclustering and FLOC Algorithm
##Delta
###Test Code for function
```{r}
set.seed(12356)
m1 <- matrix(rnorm(20),nrow=5,ncol=4)
m2 <- m1 - mean(m1)
m3 <- m1
m3[2,] <- NA
m4 <- m3 - mean(m3,na.rm=TRUE)
```
```{r}
rcdel(m4,1)
```
```{r}
NodeDel(m1,0.1)
```
Latest problem is that lack of a bicluster leads to NA in final test.  Also, looks
like coherence is not monotonic downward. [??]

```{r}
set.seed(12356)
m1 <- matrix(rnorm(20),nrow=5,ncol=4)
for (i in 1:2) for (j in 3:4) m1[i,j] <- m1[i,j] +3
NodeDel(m1,0.1)
```
### 5/13/2025
Somehow the signs look wrong.  Did not check carfully. More tomorrow
### 5/14/2025
```{r}
NodeDel(m1,1)
```

Still a problem.  Typo in book had rij=yij-aij-bij-mu, last should be a plus[??]
### 5/15/2025
```{r}
set.seed(12356)
m1 <- matrix(rnorm(20),nrow=5,ncol=4)
for (i in 1:2) for (j in 3:4) m1[i,j] <- m1[i,j] +3
```
```{r}
NodeDel(m1,1)
```
### 5/17/2025
Maybe the matrix is too small?
```{r}
options(digits=2);options(width=150)
set.seed(12356)
m1 <- matrix(rnorm(20),nrow=20,ncol=20)
for (i in 5:9) for (j in 10:15) m1[i,j] <- m1[i,j] +3
```
```{r}
set.seed(12356)
m2 <- matrix(rbinom(400,1,0.3),nrow=20,ncol=20)
for (i in 5:9) for (j in 10:15) m2[i,j] <- rbinom(1,1,0.8)
```
heatmap(m1,Rowv=NA,Colv=NA,revC=TRUE)
```{r}
NodeDel(m1,0.05)
```
gets first 4 row right, but not columns?
### 5/20/2025
Problem is that I seem to be doing calculations over the whole matrix during each iteration. 
But algorithms are not clear (to me at least)
biclust uses BCCC for Cheng & Church

Downloaded code for single node deletion from biclust
Looks somewhat like mine??
```{r}
set.seed(12356)
m1 <- nmat(50,0, 0.1,20:30,10:20,3,0.1)
```
```{r}
tmp <- bigcc(m1,1)
tmp
```
Not clear at all why this is all true.
```{r}
biclust(m1, method=BCCC(), delta = 1.0, alpha=1.5, number=1)
```
Well, it's the same answer, but not clear why???

### 5/23/2025
Rereading Kasim, not clear that my nmat normal matrix actually has what they consider 
a bicluster.
```{r}
set.seed(12356)
m1 <- nmat(50,0, 0.1,20:30,10:20,3,0.1)
m1[20:30,10:20] <- 3
```
```{r}
bc1 <- biclust(m1, method=BCCC(), delta = 1.0, alpha=1.5, number=1)
bc1
```
Nope!
Ditto for bigcc????

Looked at output of first call to single node deletion, m1 definitely has the bc
but return is input matrix m1
```{r}
set.seed(12356)
m1 <- nmat(50,0, 0.1,20:30,10:20,3,0.1)
m1[20:30,10:20] <- 3
```
```{r}
m1 <- nmat(50,0, 0.1,20:30,10:20,3,0.1)
m1[20:30,10:20] <- 3
rowscore(m1[20:30,10:20])
colscore(m1[20:30,10:20])
rowscore(m1[31:50,1:50])
bigcc(m1,0.000001)
```
One problem is that delta is too large in previous runs. But this only finds a 2x2 bicluster. 
Algorithm is just what I had, although some of my implementation is not as elegant.

### 5/24/2025
```{r}
tmp <- bigcc(m1,0.001)
```

So problem is that the residual^2 row scores are too small for the noise, i.e. they look like biclusters to the algorithm.
```{r}
set.seed(12356)
m1 <- nmat(50,3, 0.1,20:30,10:20,3,0.1)
m1[20:30,10:20] <- 3
```
Same here. They are all too close to 3.
```{r}
set.seed(12356)
m1 <- nmat(50,3,3,20:30,10:20,3,3)
m1[20:30,10:20] <- 3
bc1 <- biclust(m1, method=BCCC(), delta = 0.5, alpha=1.5, number=1)
bc1
```
This is the problem: off-biclust were small and too homogeneous to let the "real" bicluster be detected. 
Note that decreasing delta only helps up to a point.  Went from 0.1 to 0.0001 and started losing some
of the "real" bicluster.
With biclust fn, from 0.1 on down got 8x8 bicluster

So the (now) obvious point is that background has to have sufficiently high
variance compared to the bicluster(s) in order for node deletion to work.


### 5/25/2025
Now I need to test my code with this insight
```{r}
set.seed(12356)
m1 <- nmat(50,1,3,20:30,10:20,3,3)
m1[20:30,10:20] <- 3
bc1 <- biclust(m1, method=BCCC(), delta = 0.5, alpha=1.5, number=1)
bc1
```
Even with mean 1, sd=3 generates enough noise.
With 2 and 1 bicluster expands.  At sd = 0.5, whole matrix is selected
```{r}
NodeDel(m1,0.5)
```

## Multiple node deletion

# Git refresher

github name lgalway
